### 2024-10-09 manuscript/05-conclusion.qmd

# Conclusion {#sec-conclusion}
In this study, we embarked on a comprehensive analysis of pharmaceutical demand forecasting within the Ethiopian 
Pharmaceutical Supply Services (EPSS) context. By leveraging five  years of sales data for 37 key pharmaceutical 
products, we aimed to  enhance the accuracy of demand predictions and consequently  contribute to more effective supply 
chain management. Our findings  underscore the significance of accurate demand forecasting in  optimizing inventory 
levels, resource allocation, and overall supply chain operations.

The application of diverse forecasting models yielded valuable insights into their respective performance. Notably, the 
Autoregressive Integrated Moving Average (ARIMA) model emerged as a  robust choice, consistently exhibiting superior forecast accuracy  across a majority of the selected pharmaceutical products.  Additionally, the incorporation of predictor variables, such as  physical count, budget release and closure by the government,  COVID-19 impact, and campaign events, further underscored the  efficacy of the ARIMA model in accommodating external influences and delivering reliable forecasts.

The individualized assessment of forecasting models revealed specific strengths across pharmaceutical categories. ARIMA 
models with predictors demonstrated exceptional performance for pharmaceuticals including Amoxacillin 500mg capsule, Anti RhD, Artehmether + Lumefantrine, Atenolol, and others, affirming their utility in capturing complex demand patterns. 
Moreover, the study shed light on pharmaceuticals best suited for alternative models like naive, mean, and regression 
methods, contributing to a comprehensive understanding of forecasting applicability.

While these findings contribute significantly to the discourse on pharmaceutical demand forecasting, it's essential to 
acknowledge limitations. The accuracy of our forecasts is intrinsically linked to the quality and completeness of 
historical sales data, and the intricate interplay of external variables adds complexity to the forecasting landscape.

In conclusion, this study serves as a pivotal step towards enhancing pharmaceutical supply chain management through 
refined demand forecasting. The demonstrated success of ARIMA models, particularly when augmented with predictor 
variables, holds promise for fostering resilient and adaptable supply chains. As the pharmaceutical landscape continues 
to evolve, our insights encourage further exploration of advanced forecasting techniques and continuous refinement of 
strategies to bolster supply chain efficiency and ensure consistent availability of essential pharmaceuticals.

In light of the study's findings, several actionable recommendations emerge. First and foremost, the integration of the 
Autoregressive Integrated Moving Average (ARIMA) model with predictor variables stands out as a viable strategy for 
enhancing pharmaceutical demand forecasting accuracy within Ethiopian Pharmaceutical Supply Services (EPSS). This entails 
harnessing the predictive power of external factors to refine forecasting precision. To further advance forecasting 
practices, the exploration of additional predictor variables such as public health campaigns, disease outbreaks, and 
economic indicators could offer valuable insights into demand dynamics. Moreover, refining the granularity of historical 
sales data, possibly through finer temporal intervals and geographic diversity, holds potential for more nuanced and 
accurate predictions. Consideration could also be given to hybrid forecasting models that amalgamate ARIMA's short-term 
accuracy with machine learning's predictive prowess for long-term trends. Regular model evaluation, sensitivity analysis, and collaborative stakeholder engagement are recommended to ensure ongoing accuracy and alignment with changing demand patterns.

Capacitating EPSS staff through training programs in forecasting methodologies and data interpretation can empower the 
workforce to make informed decisions. Replication of this study in diverse healthcare settings and comparative analyses 
can validate the findings' applicability and adaptability. By embracing these recommendations, EPSS can fortify its 
pharmaceutical supply chain with resilient, efficient, and responsive operations, ultimately ensuring the consistent 
availability of vital pharmaceuticals.
  a