### 2024-10-09 manuscript/03-experiment.qmd

# Experiment setup {#sec-experiment}

## Data {#sec-data}

### Dataset

For this study, we utilized a dataset spanning five years (2018-2022) of sales data obtained from EPSS. Rigorous checks were conducted to ensure the consistency and completeness of the collected data. From the extensive pool of pharmaceutical products within EPSS, we strategically selected a set of 37 key pharmaceuticals, encompassing various programs and representing different classes of drugs. This deliberate selection enriches the research by providing a comprehensive representation across the pharmaceutical landscape.
We are presenting the graph for some of the pharmaceuticals sales for the last five years by using visualization for the readers to undesrtand the sales of the pharmaceuticals.

```{r}
#| include: false
library(tidyverse)
library(lubridate)
library(fpp3)
library(ggthemes)
monthly_issues_tsibble <- read_rds("../data/monthly_issues_tsibble.rds")
monthly_issues_tsibble <- monthly_issues_tsibble |> filter_index("2017 Aug" ~ "2022 Jun")

knitr::opts_chunk$set(fig.align = 'center', out.width = "90%")

monthly_issues_tsibble |> as_tibble() |> select(month) -> predictors

write_csv(predictors, "../data/predictors.csv")
getwd()
```

```{r}
#| include: false
items <- unique(monthly_issues_tsibble$item)
issue_plot_monthly <- function(my_item) {
  ggplot(data = monthly_issues_tsibble |> filter(item == my_item), mapping = aes(x = month, y = quantity_product_issued))+
    geom_point(size=1)+
    geom_line()+
    theme_few()+
    labs(title=my_item)+
    scale_x_yearmonth(date_breaks = "2 month")+
    theme(axis.text.x = element_text(angle=90))
}

my_plot <- list()
for (i in (1:40)) {
  my_plot[[i]] <- issue_plot_monthly(items[i])
}
```

## Visualising monthly time series

```{r}
my_plot[[1]]
```

```{r}
my_plot[[2]]
```

```{r}
my_plot[[4]]
```

```{r}
my_plot[[6]]
```
```{r}
my_plot[[9]]
```

```{r}
my_plot[[11]]
```
```{r}
my_plot[[12]]
```
```{r}
my_plot[[15]]
```
```{r}
my_plot[[17]]
```
```{r}
my_plot[[19]]
```
```{r}
my_plot[[21]]
```
```{r}
my_plot[[25]]
```
```{r}
my_plot[[26]]
```
```{r}
my_plot[[29]]
```
```{r}
my_plot[[30]]
```
```{r}
my_plot[[32]]
```
```{r}
my_plot[[36]]
```
```{r}
my_plot[[37]]
```

# 2. 2. Data Splitting and Preparation

The primary objective of this study is to forecast pharmaceutical demand within EPSS for the year 2023. To achieve this, the historical sales data for the chosen pharmaceuticals