### 2024-10-10 manuscript/02-literature.qmd
Revised literature review

# Research background {#sec-lit}

Accurate sales forecasting is crucial in the pharmaceutical industry, where it directly impacts profit maximization, cost minimization, and the ability to respond to market changes. Forecasting in healthcare not only influences clinical decisions but also plays a pivotal role in supply chain management, ensuring that the right drugs are available when needed. The challenge lies in balancing customer demand with inventory costs, which is particularly complex in pharmaceuticals due to factors like short shelf life and quality constraints [@gupta2000mid, @makridakis2020forecasting].

In this review, a range of forecasting methods has been applied, from traditional statistical models to advanced machine learning techniques, with many studies utilizing a combination of simple and advanced methods to enhance accuracy and reliability[@rathipriya2023demand,@merkuryeva2019demand,@nikolopoulos2016forecasting,@van2021using,@khalil2014intelligent,@zhu2021demand,@anusha2014demand,@burinskiene2022forecasting,@bon2017optimization,@sousa2019statistical,@zahra2019forecasting]. Some studies, such as those by @newberne2006holt, and @restyana2021analysis, applied traditional forecasting methods like the Holt-Winters method, Single Moving Average (SMA), and Single Exponential Smoothing (SES), which are simpler but effective for certain applications. In contrast, other researchers explored more advanced approaches, including hybrid models and machine learning techniques, to address complex forecasting challenges, as seen in the studies by @siddiqui2022hybrid, @de2021lead, @kim2015demand,@candan2014demand, and @ribeiro2017improving].

Early research in pharmaceutical forecasting primarily focused on traditional statistical methods. These methods, including ARIMA, exponential smoothing, and moving averages, were widely used due to their simplicity and effectiveness in relatively stable environments [@zahra2019forecasting]. These models typically assess forecast accuracy using metrics like Mean Absolute Percentage Error (MAPE), Mean Squared Error (MSE), and Root Mean Squared Error (RMSE). For example, @newberne2006holt applied the Holt-Winters method to forecast healthcare data, focusing on prescription trends. Their findings highlighted the method's utility in short-term planning and resource management, validated through metrics like MAPE, MAD, and MSD. Similarly, @restyana2021analysis compared Single Moving Average (SMA) and Single Exponential Smoothing (SES) in drug demand forecasting, concluding that SES provided more accurate results, as indicated by lower MAD and MSE values.

Recognizing the limitations of traditional models, particularly in capturing the complexities of pharmaceutical demand (e.g., seasonality, external influences), researchers began developing hybrid models. These models combine linear methods like ARIMA with nonlinear techniques such as neural networks, aiming to improve forecasting accuracy by addressing diverse data patterns. @khalil2014intelligent introduced a hybrid model that integrated ARIMA with neural networks. Their study demonstrated that this approach significantly outperformed traditional models, especially in scenarios with limited historical data. The hybrid model's effectiveness was confirmed by improvements in key performance indicators like MAPE and RMSE.

The pharmaceutical industry's increasing complexity and the availability of large datasets have led to the adoption of advanced machine learning techniques. These methods, such as Support Vector Regression (SVR), Random Forest (RF), and Long Short-Term Memory (LSTM) networks, offer greater flexibility and predictive power.@van2021using, showed that incorporating downstream information into machine learning models, like LASSO and SVR, greatly enhances forecast accuracy. Their study on multi-echelon supply chains revealed that advanced models, particularly those integrating external variables, consistently produced lower forecast errors, as measured by AvgRelRMSE. Similarly, @rathipriya2023demand compared shallow neural networks and deep learning models for drug demand forecasting. The study found that shallow models outperformed deep learning models for most drug categories, while ARIMA was more effective for the remaining categories. This suggests that no single model is universally optimal, and model selection should be context-specific.

More sophisticated models, such as Neuro-Fuzzy Systems, have emerged as powerful tools for pharmaceutical forecasting. These systems combine the learning capabilities of neural networks with the reasoning abilities of fuzzy logic, offering a balanced approach that incorporates both empirical data and expert knowledge. Deep learning models, while powerful, present challenges such as the need for large datasets and substantial computational resources. Despite these challenges, models like LSTM have shown potential in capturing long-term dependencies in time series data, as demonstrated by @sousa2019statistical in their study on drug distribution in the Brazilian Public Health System. @candan2014demand employed an Adaptive Neuro-Fuzzy Inference System (ANFIS) to forecast pharmaceutical demand. Their study highlighted the system's ability to capture complex patterns in the data, resulting in highly accurate forecasts. The effectiveness of this approach was further validated through statistical tests, such as paired T-tests and mean difference analysis, which confirmed its superiority over traditional methods.

Despite the advancements in forecasting methods, several gaps remain, particularly in addressing uncertainty and incorporating domain knowledge into the modeling process. Most existing studies focus on point forecasts, which may not fully capture the range of possible outcomes in pharmaceutical demand. This limitation is critical, given the high stakes of inventory management in this industry. Moreover, there is a lack of research on the impact of integrating domain knowledge—such as insights from experienced professionals—into forecasting models. This integration could potentially improve model accuracy and reliability, especially in complex and rapidly changing environments. In summary, the literature on pharmaceutical forecasting has evolved from traditional statistical models to advanced machine learning techniques, each offering unique insights and capabilities. However, there is still room for improvement, particularly in addressing uncertainty and incorporating domain knowledge. Future research should focus on developing probabilistic forecasting models and exploring ways to integrate expert insights into the forecasting process. These advancements could significantly enhance decision-making and operational efficiency in the pharmaceutical supply chain.

# Table 1: Summary of some studies on forecasting in pharmaceutical supply chain

# How the domain knowledge were collected

To gather domain knowledge, we contacted experts from the Ethiopian Pharmaceuticals Supply Service (EPSS) who work in the distribution unit. Five experts with over 10 years of experience in managing the distribution of the pharmaceuticals in EPSS agreed to participate. The researchers, along with these experts, examined the electronic bin cards of selected pharmaceuticals. A bin card is a physical or electronic record used to track the inventory levels of individual items in a storage location, such as a warehouse or stockroom. The term 'bin' refers to the specific storage location for an item. Information typically recorded on a bin card includes the Item ID, unit of measure, product description, account details, date, references, beginning balance, manufacturer, batch number, expiry date, quantity received, quantity issued, and balance. In summary, a bin card is a fundamental tool in inventory management, especially in environments where physical stock tracking is essential. It provides a straightforward and immediate way to monitor item quantities, helping to ensure effective inventory management.

We meticulously traced sales data for the past five years, triangulating graphs with bin card records to identify predictors influencing demand for each product. The expert uses their domain knowledge and experience to identify about the distribution charachterstics of the specific products and sometimes when they assume that they do not have enough history or do not remember about that specific products they call to other experts who remember specific issues and decide whetrere the specific predictos affect the distribution and if all of them agree the include it as predcitors. The data collection process took more than a month. We downloaded and analyzed over 100 bin cards per product. Dates on the bin cards were converted from the Ethiopian calendar to the Gregorian calendar for accuracy. Transactions due to storage or warehouse changes within EPSS were excluded. After cleaning the data,and discussing with the experts  we identified the following key predictors affecting distribution and sales: stock replenishment, stock outs, COVID-19, fiscal year counting, conflict, and malaria seasonality.

•	Stock replenishment:  refers to the process of restocking or refilling inventory to ensure that there are sufficient quantities of products or materials available to meet demand. Whenever there was stock replenishment at the central EPSS, sales and distribution to hubs and health facilities increased. This increase was attributed to the need to restock depleted inventories and the push from central EPSS to manage space constraints. 
•	Stockouts:  occur when an inventory item is completely depleted, meaning there is no available stock to meet demand. This can happen when demand exceeds the available supply, or when there is a delay or failure in the replenishment process. Resulted in no transactions and a decline in sales.

•	COVID-19: COVID-19, caused by the SARS-CoV-2 virus, had a profound impact on Ethiopia's pharmaceutical supply chain. The pandemic disrupted global manufacturing and supply, leading to shortages of essential medicines and raw materials, Ethiopia, like many other countries, relies heavily on imported pharmaceutical products. With the pandemic causing lockdowns and restrictions in manufacturing hubs like India and China, the supply of essential medicines and raw materials to Ethiopia was significantly disrupted. The impact of COVID-19 on sales was significant. After the announcement of COVID-19 in Ethiopia, the central EPSS pushed a six-month supply of products to hubs to prevent stock outs. This push was from March to August 2020. During this period, if products were not stocked out, sales continued to hubs and health facilities.

•	Physical inventory: refers to the process of manually counting and verifying the actual quantities of pharmaceutical products available in stock at a specific location. The process is critical for maintaining the accuracy of inventory records, ensuring that medicines are available when needed, and preventing stockouts or overstocking.  Physical inventory counting periods also influenced sales. Stores closed during these periods, halting transactions. We observed increased sales before inventory counting periods, as hubs and facilities stocked up. July and August were identified as physical counting periods each year.

•	Internal displacement: Occurs when people are forced to flee their homes but remain within their country's borders, typically due to conflicts, violence, or natural disasters. Conflicts, whether ethnic, political, or resource-driven, can lead to widespread instability, forcing large populations to move and creating humanitarian crises. Conflicts often lead to the destruction of infrastructure, including roads, bridges, and warehouses, making it difficult to transport pharmaceutical products to affected areas. This disruption can cause significant delays or prevent the delivery of essential medicines altogether. Internal displacements and conflicts in Ethiopia over the past five years also affected sales and distribution. We traced periods of instability and included them as predictors. The identified conflict periods were September 2017 to March 2018, February 2019 to June 2019, and November 2020 to July 2022.

•	Malaria seasonality: Refers to the predictable patterns and fluctuations in malaria incidence throughout the year, typically influenced by climate and environmental conditions. In many regions, malaria transmission peaks during and shortly after the rainy season, when conditions such as stagnant water pools create ideal breeding sites for the Anopheles mosquitoes that transmit the disease. Conversely, malaria cases often decline during the dry season when mosquito breeding sites are reduced. During peak malaria seasons, there is a significant surge in the demand for antimalarial drugs and other related treatments. Malaria seasonality was another significant predictor. Certain pharmaceuticals, like Artemether + Lumefanthrine and Rapid Diagnostic Test kits, were affected by malaria outbreaks. We identified epidemic periods affecting sales: September to December 2017, March to May 2018, September to December 2018, March to May 2019, September to December 2019, March to May 2020, September to December 2020, March to May 2021, September to December 2021, and March to May 2022.

The comprehensive analysis and inclusion of these predictors ensured a robust model for forecasting pharmaceutical sales and distribution, leveraging expert domain knowledge and extensive historical data analysis.