### 2024-10-12 manuscript/01-introduction.qmd
Introduction revised according to Prof Bahman comments

---
bibliography: references.bib
---
# Introduction {#sec-intro}

Universal access to essential medicines is a fundamental goal of effective healthcare systems, yet ensuring equitable availability continues to be a critical global challenge [@quick2003essential; @world2004annual]. Many countries, particularly in Africa and Asia, face significant barriers to accessing these medications, often due to high costs and inefficient supply chain management [@world2004medicines]. A survey in eight sub-Saharan African countries revealed concerningly low availability of essential medicines, with mean availability ranging from 22% to 40% for women’s priority medicines and from 28% to 57% for children’s medicines [@droti2019poor]. However, drug shortages are not limited to low-income regions; they also occur in developed areas like the United States and Europe, where they disrupt healthcare delivery and compromise the quality of patient care [@fox2003managing; @kaakeh2011impact; @johnson2011drug; @huys2013european; @le2011prevalence].

The consequences of drug shortages extend beyond immediate healthcare delivery, impacting healthcare costs, treatment quality, and patient safety [@alspach2012drug; @kaakeh2011impact; @baumer2004national]. Such shortages can result in treatment delays, suboptimal care due to the unavailability of preferred therapies, increased reliance on costly secondary markets, and serious patient outcomes, including complications, longer hospital stays, and even deaths [@alspach2012drug]. Addressing these challenges requires a resilient pharmaceutical supply chain that can ensure the timely availability of medicines. In this context, accurate demand forecasting becomes critical for effective procurement and inventory management [@subramanian2021effective].

Demand forecasting in pharmaceutical supply chains is a complex endeavor, influenced by factors such as data quality, supply chain positioning, and external drivers, including market trends and regulatory changes [@schneider2010pharmaceutical; @hyndman2018forecasting]. While vast amounts of data are often collected for transactional purposes—such as sales—many forecasting approaches overlook the significance of domain knowledge, which is as crucial as quantitative data for understanding the context and making more informed predictions. Domain expertise helps identify the impact of nuanced factors, such as clinical guidelines, disease outbreaks, and local policies, that are not always evident from sales data alone.

Combining statistical models with domain insights offers a more refined forecasting process, as evidenced by research that highlights the value of judgmental adjustments to enhance forecast accuracy [@fildes2007against; @taylor2022artificial; @soyiri2013overview]. Domain knowledge enables analysts to incorporate relevant factors, such as market conditions, policy changes, and supply chain disruptions, into the forecasting model. It aids in guiding feature selection, parameter tuning, and the interpretation of complex data patterns, ultimately leading to more accurate and actionable forecasts [@sinha2008incorporating; @dash2022review]. This integration is particularly beneficial in fields like healthcare, where expert insights can align machine learning algorithms with real-world applications and ethical considerations [@dash2022review].

Despite the considerable body of research on integrating domain knowledge with statistical techniques, relatively few studies focus specifically on the unique requirements of pharmaceutical forecasting. This study aims to bridge that gap by employing advanced machine learning models combined with domain expertise to address the distinctive characteristics of Ethiopia’s pharmaceutical sector. The challenges in this sector, such as diverse product categories, limited data from service delivery points, communication issues, long lead times, forecast inaccuracies, and complex policies, necessitate tailored forecasting approaches [@bilal2024challenges; @boche2022procurement]. Accurate demand forecasting in Ethiopia is essential for optimizing resource allocation, streamlining supply chain operations, and shaping effective healthcare policies [@rostami2022forecasting].

This research will forecast the demand for selected pharmaceuticals in the Ethiopian Pharmaceutical Supply Service (EPSS) over the next six months using various methods, including dynamic regression ARIMA, multiple regression, neural networks, machine learning, and time-series GPT models. We will explore whether incorporating domain knowledge can improve forecast accuracy by integrating predictors derived from historical events over the past five years, such as stock replenishment cycles, stockouts, fiscal calendar variations, COVID-19, conflicts, and seasonal trends like malaria outbreaks.

Our contributions include not only generating point forecasts but also employing probabilistic forecasting techniques that capture the range of potential outcomes with associated probabilities, offering a more comprehensive view of demand uncertainty than traditional methods. Additionally, we identify key factors influencing pharmaceutical demand and incorporate these into our forecasting models to better reflect the complexities of the Ethiopian context. Model performance will be evaluated using metrics such as Root Mean Squared Error (RMSE), Mean Absolute Error (MAE), and probabilistic measures like the Continuous Ranked Probability Score (CRPS) and Winkler score.

The remainder of this paper is structured as follows: Section 2 reviews the literature and identifies gaps to position our work; Section 3 outlines the experimental design, including data sources, forecasting methods, and evaluation criteria; Section 4 presents the results and analysis; and Section 5 concludes with a summary of findings and future research directions.