### 2024-10-24 manuscript/02-literature.qmd

# Research background {#sec-lit}

Accurate sales forecasting is crucial in the pharmaceutical industry, where it directly impacts profit maximization, cost minimization, and the ability to respond to market changes. Forecasting in healthcare not only influences clinical decisions but also plays a pivotal role in supply chain management, ensuring that the right drugs are available when needed. The challenge lies in balancing customer demand with inventory costs, which is particularly complex in pharmaceuticals due to factors like short shelf life and quality constraints [@gupta2000mid, @makridakis2020forecasting].

In this review, a range of forecasting methods has been applied, from traditional statistical models to advanced machine learning techniques, with many studies using a combination of simple and advanced methods to enhance accuracy [@nikolopoulos2016forecasting,@zhu2021demand,@anusha2014demand]. Some studies, such as those by @newberne2006holt, and @restyana2021analysis, applied traditional forecasting methods like the Holt-Winters method, Single Moving Average (SMA), and Single Exponential Smoothing (SES), which are simpler but effective for certain applications. In contrast, other researchers explored more advanced approaches, including hybrid models and machine learning techniques, to address complex forecasting challenges, as seen in the studies by @siddiqui2022hybrid, @de2021lead, @kim2015demand,@candan2014demand, and @ribeiro2017improving].

Early research in pharmaceutical forecasting primarily focused on traditional statistical methods. These methods, including ARIMA, exponential smoothing, and moving averages, were widely used due to their simplicity and effectiveness in relatively stable environments [@zahra2019forecasting]. These models typically assess forecast accuracy using metrics like Mean Absolute Percentage Error (MAPE), Mean Squared Error (MSE), and Root Mean Squared Error (RMSE). For example, @newberne2006holt applied the Holt-Winters method to forecast healthcare data, focusing on prescription trends. Their findings highlighted the method's utility in short-term planning and resource management, validated through metrics like MAPE, MAD, and MSD. Similarly, @restyana2021analysis compared Single Moving Average (SMA) and Single Exponential Smoothing (SES) in drug demand forecasting, concluding that SES provided more accurate results, as indicated by lower MAD and MSE values.

Recognizing the limitations of traditional models, particularly in capturing the complexities of pharmaceutical demand (e.g., seasonality, external influences), researchers began developing hybrid models. These models combine linear methods like ARIMA with nonlinear techniques such as neural networks, aiming to improve forecasting accuracy by addressing diverse data patterns. @khalil2014intelligent introduced a hybrid model that integrated ARIMA with neural networks. Their study demonstrated that this approach significantly outperformed traditional models, especially in scenarios with limited historical data. The hybrid model's effectiveness was confirmed by improvements in key performance indicators like MAPE and RMSE.

The pharmaceutical industry's increasing complexity and the availability of large datasets have led to the adoption of advanced machine learning techniques. These methods, such as Support Vector Regression (SVR), Random Forest (RF), and Long Short-Term Memory (LSTM) networks, offer greater flexibility and predictive power.@van2021using, showed that incorporating downstream information into machine learning models, like LASSO and SVR, greatly enhances forecast accuracy. Their study on multi-echelon supply chains revealed that advanced models, particularly those integrating external variables, consistently produced lower forecast errors, as measured by AvgRelRMSE. Similarly, @rathipriya2023demand compared shallow neural networks and deep learning models for drug demand forecasting. The study found that shallow models outperformed deep learning models for most drug categories, while ARIMA was more effective for the remaining categories. This suggests that no single model is universally optimal, and model selection should be context-specific.

More sophisticated models, such as Neuro-Fuzzy Systems, have emerged as powerful tools for pharmaceutical forecasting. These systems combine the learning capabilities of neural networks with the reasoning abilities of fuzzy logic, offering a balanced approach that incorporates both empirical data and expert knowledge. Deep learning models, while powerful, present challenges such as the need for large datasets and substantial computational resources. Despite these challenges, models like LSTM have shown potential in capturing long-term dependencies in time series data, as demonstrated by @sousa2019statistical in their study on drug distribution in the Brazilian Public Health System. @candan2014demand employed an Adaptive Neuro-Fuzzy Inference System (ANFIS) to forecast pharmaceutical demand. Their study highlighted the system's ability to capture complex patterns in the data, resulting in highly accurate forecasts. The effectiveness of this approach was further validated through statistical tests, such as paired T-tests and mean difference analysis, which confirmed its superiority over traditional methods.

Despite the advancements in forecasting methods, several gaps remain, particularly in addressing uncertainty and incorporating domain knowledge into the modeling process. Most existing studies focus on point forecasts, which may not fully capture the range of possible outcomes in pharmaceutical demand. This limitation is critical, given the high stakes of inventory management in this industry. Moreover, there is a lack of research on the impact of integrating domain knowledge—such as insights from experienced professionals—into forecasting models. This integration could potentially improve model accuracy and reliability, especially in complex and rapidly changing environments. In summary, the literature on pharmaceutical forecasting has evolved from traditional statistical models to advanced machine learning techniques, each offering unique insights and capabilities. However, there is still room for improvement, particularly in addressing uncertainty and incorporating domain knowledge. Future research should focus on developing probabilistic forecasting models and exploring ways to integrate expert insights into the forecasting process. These advancements could significantly enhance decision-making and operational efficiency in the pharmaceutical supply chain.

@tbl-literature provides a summary of key studies in the literature on forecasting for pharmaceutical products. We identify several limitations both in the existing research and in practice, which highlight important gaps that motivate our current study. These gaps are summarized as follows:

1. Despite significant investment in data collection technologies and logistics management systems, these systems primarily collect transactional data, such as consumption. However, no data is collected on events that influence consumption variability—information that would be crucial for building reliable forecasting models. No studies to date have explored how domain expertise could enhance pharmaceutical demand forecasting.

2. Current research on forecasting for pharmaceutical supply chain predominantly focuses on generating point forecasts. There is a lack of studies that consider the entire forecast distribution of monthly consumption, which would better capture the uncertainty of future demand and provide a valuable risk management tool for decision-makers.

3. Reproducibility remains a major challenge in this field. It is often difficult for readers to reproduce previous studies without direct assistance from the authors, limiting the practical application and validation of existing research.

```{r}
#| label: tbl-literature
#| echo: false
#| cache: false
#| tbl-cap: "Summary of some studies on forecasting in pharmaceutical supply chain"
library(kableExtra)
readxl::read_xlsx("Literature_review_table.xlsx") |>
  knitr::kable(booktabs = T, linesep = "") |>
  kable_styling(latex_options = "scale_down", font_size = 11) |>
  column_spec(3, "6em") |>
  column_spec(5, "4em") |>
  column_spec(6, "10em") |>
  column_spec(7, "8em") |>
  landscape()
```