### 2025-05-26 diagram/diagram.qmd
revision submitted

```{=latex}
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[scale=0.85, transform shape, node distance=1.3cm and 2cm]

% Node style
\tikzset{
  process/.style={
    rectangle,
    draw=black,
    thick,
    text width=4.5cm,
    align=center,
    minimum height=1.4cm
  },
  arrow/.style={
    thick,
    ->,
    >=Stealth
  }
}

% Nodes
\node (step1) [process] {1. Expert Identification (with EPSS)};
\node (step2) [process, below=of step1] {2. Onboarding \& Orientation};
\node (step3) [process, below=of step2] {3. Workshop: Collaborative Data Review \& Joint Identification of Anomalies};

\node (step4) [process, right=of step1] {4. Cross Validation Using Bin Cards \& Inventory Records};
\node (step5) [process, below=of step4] {5. Structured Expert Elicitation \& Contextual Clarification};

\node (step6) [process, right=of step4] {6. Expert Elicitation of Predictor Relevance};
\node (step7) [process, below=of step6] {7. Consensus-Based Selection of Predictors};
\node (step8) [process, below=of step7] {8. Variable Encoding \& Time Alignment};

\node (step9) [process, below=of step8, yshift=-0.3cm] {9. Integration into Forecasting Models};

% Arrows
\draw [arrow] (step1) -- (step2);
\draw [arrow] (step2) -- (step3);

\path (step3.east) -- +(0.6,0) coordinate (s3stub);
\draw [arrow] (step3.east) -- (s3stub) |- (step4.west);

\draw [arrow] (step4) -- (step5);

\path (step5.east) -- +(0.6,0) coordinate (s5stub);
\draw [arrow] (step5.east) -- (s5stub) |- (step6.west);

\draw [arrow] (step6) -- (step7);
\draw [arrow] (step7) -- (step8);
\draw [arrow] (step8) -- (step9);

% Group boxes
\node[draw=black, thick, dashed, inner sep=0.3cm, rounded corners, fit=(step1)(step2)(step3), label={[black]90:Expert Engagement}] {};
\node[draw=black, thick, dashed, inner sep=0.3cm, rounded corners, fit=(step4)(step5), label={[black]90:Validation}] {};
\node[draw=black, thick, dashed, inner sep=0.3cm, rounded corners, fit=(step6)(step7)(step8), label={[black]90:Feature Construction}] {};
\node[draw=black, thick, dashed, inner sep=0.3cm, rounded corners, fit=(step9), label={[black]90:Model Integration}] {};

\end{tikzpicture}
\end{figure}

```