### 2025-05-26 manuscript/05-conclusion.qmd

# Conclusion and future reserach {#sec-conclusion}

In this study, we conducted an extensive analysis of pharmaceutical demand forecasting within the EPSS. Using five years of consumption data for 33 key pharmaceutical products, along with additional information gathered through collaboration with domain experts at EPSS, our goal is to contribute to enhancing the demand forecasting process and emphasize the importance of integrating domain knowledge into model building. This step is critical, as the phase of understanding the data and incorporating valuable contextual information is often overlooked. Too frequently, forecasting efforts rely solely on consumption data, leading to building models with little relevance to reality. Our approach highlights the necessity of including domain insights to construct more accurate and effective forecasting models.

Our findings highlight the importance of collecting and incorporating domain knowledge when building forecasting models. We evaluated both point and probabilistic forecasts using a range of models, from simple univariate approaches like ARIMA to complex models such as LSTM. We demonstrated that while the use of complex models may improve forecast accuracy in the pharmaceutical supply context, this comes at the cost of significantly increased computational time—an important consideration for resource-constrained settings in low- and middle-income countries (LMICs). Additionally, our research indicates that newly developed foundational time series forecasting models hold promise, particularly in settings with limited computational infrastructure and constrained access to advanced analytical expertise, as often seen in pharmaceutical supply chains in developing countries.

To advance forecasting practices further, exploring additional predictor variables—such as public health campaigns, disease outbreaks, and economic indicators—could provide valuable insights into demand dynamics. Improving the granularity of historical consumption data, through finer temporal intervals and geographic differentiation, also holds potential for more precise predictions. Additionally, hybrid forecasting models that combine judgmental point forecasts with probabilistic machine learning models could enhance predictive performance. Empowering EPSS staff through training in foundational time series models, data interpretation, and collection will enable informed decision-making and strengthen workforce capabilities. Replicating this study across diverse healthcare settings and conducting comparative analyses would help validate the adaptability and applicability of the findings. Additionally, it is recommended that pharmaceutical supply services systematically collect and maintain detailed records of events that may influence consumption, alongside consumption data. This practice is essential for refining demand models and enhancing forecast accuracy.

## Practical challenges and limitations

Despite dedicated efforts to engage with domain experts to better understand the data—and spending significant time over two weeks collaborating with them—we found that the process of interpreting data through domain knowledge is complex and presents unique challenges. These challenges include defining what constitutes domain knowledge and determining how best to incorporate it to enhance model reliability and accuracy. Below, we summarize some of these key challenges. One issue involved capturing accurate information from bin cards and expert review, particularly during periods of disruption like COVID-19 and conflicts. For instance, during the COVID-19 pandemic, experts noted a decrease in demand due to travel restrictions. However, despite the reduced demand, there were still significant distributions of pharmaceuticals from the EPSS central to various hubs and health facilities. This was because, in response to anticipated shortages, a political decision was made to push products to these facilities before the travel restrictions took full effect. The rationale was that if the facilities remained closed, patients would have no access to medications, necessitating the preemptive stockpiling of essential pharmaceuticals. This scenario illustrates how policy decisions can significantly impact supply chain data, making it challenging to model distribution patterns accurately. Similarly, experts indicated that conflicts would disrupt the distribution of pharmaceuticals. While conflict zones did indeed hinder transportation, distribution still occurred whenever roads were temporarily opened, even amidst on-going conflict. This created inconsistencies in the data, as periods of halted distribution were followed by rapid replenishment once access was restored. Such fluctuations add complexity to modelling the supply chain from the central EPSS to hubs and health facilities. Other challenges involved seasonal or periodic activities, such as the fiscal year-end stock counting. During this time, EPSS temporarily halts transactions to conduct a full inventory count, which can take anywhere from one to two months, depending on the circumstances. This inconsistency in the duration of inventory counts adds a layer of unpredictability to supply chain modelling.

## Future reserach

Following the current research, several promising areas could further advance knowledge and practical applications in this area:

- Future research could focus on developing intelligent systems that incorporate domain-specific knowledge into model building. This would involve creating algorithms capable of identifying relevant domain insights and integrating them effectively into model training processes. Such tools would bridge the gap between purely data-driven approaches and expert-knowledge-enhanced modeling, leading to more robust and contextually informed forecasts.

- Large Language Models (LLMs) for time series forecasting, like the one included in this study, offer transformative opportunities for developing countries by making advanced forecasting methods accessible, even in contexts with limited expertise or infrastructure. However, future research should explore their applicability, advantages, limitations, and performance relative to traditional and hybrid forecasting methods.

- \textcolor{blue}{Disruptions such as the COVID-19 pandemic and armed conflicts may alter consumption patterns in time series data, introducing irregularities that challenge the assumptions of traditional forecasting methods. In such contexts, it becomes important not only to model observed consumption but also to account for censored demand—i.e., the unmet need that would have been fulfilled under normal conditions and is essential for effective replenishment planning. Future research could focus on developing and evaluating forecasting strategies that are specifically designed to handle disrupted time series and estimate censored demand. In particular, methods that enhance the robustness of probabilistic forecasts in the presence of structural breaks or exogenous shocks are especially relevant. One promising direction is also the development of causal forecasting models that incorporate policy changes, epidemic dynamics, and socio-economic factors—an approach that may be particularly well suited to the realities of pharmaceutical supply chains in low- and middle-income countries (LMICs).}

- \textcolor{blue}{Future research should explore the integration of probabilistic forecasting into inventory management policies, with a focus on evaluating the operational impact of forecast uncertainty. In addition, there is substantial potential in leveraging more granular data—such as daily or weekly consumption patterns—and examining forecasts at hierarchical levels, including individual sites or health facilities. Investigating how fine-grained temporal and spatial forecasts influence decision-making in inventory control could provide valuable insights into reducing stockouts, minimizing holding costs, and decreasing pharmaceutical waste.}